Component response rate variation underlies stability in large complex systems

### 2019-02-22 notebook/ms.tex
First pass, fix some notation

The stability of a complex system generally decreases with increasing
system size and interconnectivity, a counterintuitive result of
widespread importance across the physical, life, and social sciences.
Despite recent interest in the relationship between system properties
and stability, the effect of variation in the response rate of
individual system components remains unconsidered. Here I vary the
component response rates (\(\boldsymbol{\gamma}\)) of randomly generated
complex systems. I show that when component response rates vary, the
potential for system stability is markedly increased. Variation in
\(\boldsymbol{\gamma}\) becomes increasingly important as system size
increases, such that the largest stable complex systems would be
unstable if not for \(\boldsymbol{Var(\gamma)}\). My results reveal a
previously unconsidered aspect of system stability that is likely to be
pervasive across all complex systems.

@@

May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
stability criterion \(\sigma\sqrt{SC} < 1\) assumes that individual
components respond to perturbations of the system at the same rate
(\(\gamma\)), but this is highly unlikely in any complex system. In
ecological communities, for example, the rate at which population
density changes following perturbation will depend on the generation
time of organisms, which might vary by orders of magnitude among
species. Species with short generation times will respond quickly (high
\(\gamma\)) to perturbations relative to species with long generation
times (low \(\gamma\)). Similarly, the speed at which individual banks
respond to perturbations in financial networks, or individuals or
institutions respond to perturbations in complex social networks, is
likely to vary. The effect of such variance on stability has not been
investigated in complex systems theory. Intuitively, variation in
\(\gamma\) might be expected to decrease system stability by introducing
a new source of variation into the system and thereby increasing
\(\sigma\). Here I show why, despite higher \(\sigma\), complex systems
in which \(\gamma\) varies are actually more likely to be stable,
especially when \(S\) is high.

@@

\textbf{Component response rates of random complex systems}. Rows in
\(\mathbf{M}\) define how a given component \(i\) is affected by other
components of the system, meaning that the rate of component response
time can be modelled by multiplying all row elements by a real scalar
value
\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
The distribution of \(\gamma\) over \(S\) components thereby models the
distribution of component response rates. An instructive example
compares one \(\mathbf{M}\) where \(\gamma_{i} = 1\) for all \(i\) in
\(S\) to the same \(\mathbf{M}\) when half of \(\gamma_{i} = 1.95\) and
half of \(\gamma_{i} = 0.05\). This models one system in which
\(\gamma\) is invariant and one in which \(\gamma\) varies, but systems
are otherwise identical (note \(E[\gamma_{i}] = 1\) in both cases). I
assume \(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\); diagonal
elements are set to \(-1\) and non-zero off-diagonal elements are drawn
randomly from \(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied
by \(\gamma_{i}\) to generate \(\mathbf{M}\). When \(\gamma_{i} = 1\),
eigenvalues of \(\mathbf{M}\) are distributed uniformly within a circle
centred at (\(-1, 0\)) with a radius of 1.265 (Fig. 1a). Hence, the real
components of eigenvalues are highly unlikely to all be negative when
all \(\gamma_{i} = 1\). But when \(\gamma_{i}\) values are separated
into two groups, eigenvalues are no longer uniformly distributed (Fig.
1b). Instead, two distinct clusters of eigenvalues appear (red circles
in Fig. 1b), one centred at (\(-1.95, 0\)) and the other centred at
(\(-0.05, 0\)). The former has a large radius, but the real components
have shifted to the left (in comparison to when \(\gamma = 1\)), while
the latter cluster has a smaller radius, but real components have
shifted to the right; all \(\Re({\lambda}) < 0\). Overall, for 1 million
randomly assembled \(\mathbf{M}\), this division between slow and fast
component response rates results in more stable systems: 1 stable given
\(\gamma = 1\) versus 32 stable given \(\gamma = \{1.95, 0.5\}\).

Higher stability in systems with variation in \(\gamma\) can be observed
by sampling \(\gamma_{i}\) values from various distributions. I focus on
a uniform distribution where \(\gamma \sim \mathcal{U}(0, 2)\) (see
Supplementary Information for other distributions, which give similar
results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig. 1b),
\(E[\gamma] = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
comparison of \(\mathbf{M}\) before and after the addition of variation
in component response rate. Figure 2 shows a comparison of eigenvalue
distributions given \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\). As
expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}}, when
\(\gamma = 1\), eigenvalues are distributed uniformly in a circle
centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 2.828.
Uniform variation in \(\gamma\) leads to a non-uniform distribution of
eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}--\protect\hyperlink{ref-Stone2017}{18}},
some of which are clustered locally near the centre of the distribution,
but others of which are spread outside the former radius of 2.828 (Fig
2b). The clustering and spreading of eigenvalues introduced by
\(Var(\gamma)\) can destabilise previously stable systems or stabilise
systems that are otherwise unstable. But where systems are otherwise too
complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
can often lead to stability above
May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
threshold \(\sigma\sqrt{SC} < 1\).

\textbf{Simulation of random \(\mathbf{M}\) across \(\mathbf{S}\)}. To
investigate the effect of \(Var(\gamma)\) on system stability, I
simulated random \(\mathbf{M}\) matrices at \(\sigma = 0.4\) and
\(C = 1\) across \(S = \{2, 3, ..., 49, 50\}\). One million
\(\mathbf{M}\) were simulated for each \(S\), and the stability of
\(\mathbf{M}\) was assessed given \(\gamma = 1\) versus
\(\gamma \sim \mathcal{U}(0, 2)\). For all \(S > 10\), I found that the
number of stable random systems was higher given \(Var(\gamma)\) than
when \(\gamma = 1\) (Fig. 3; see Supplementary Information for full
table of results), and that the difference between the probabilities of
observing a stable system increased with an increase in \(S\); i.e., the
potential for \(Var(\gamma)\) to drive stability increased with system
complexity. For the highest values of \(S\), nearly all systems that
were stable given \(Var(\gamma)\) would not have been stable given
\(\gamma = 1\).

\textbf{Targeted manipulation of \(\mathbf{\gamma}\)}. To further
investigate the potential of \(Var(\gamma)\) to be stabilising, I used a
genetic algorithm. Genetic algorithms are heuristic tools that mimic
evolution by natural selection, and are useful when the space of
potential solutions (in this case, possible combinations of \(\gamma\)
values leading to stability in a large complex system) is too large to
search
exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}.
Generations of selection on \(\gamma\) value combinations to minimise
\(\max\left(\Re(\lambda)\right)\) demonstrated the potential for
\(Var(\gamma)\) to increase system stability. Across
\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
resulted in stable systems with probabilities that were up to four
orders of magnitude higher than when \(\gamma = 1\) (Fig. 4), meaning
that stability could often be achieved by manipulating \(S\) \(\gamma\)
values rather than \(S \times S\) \(\mathbf{M}\) elements.

\textbf{System feasibility given \(\mathbf{Var(\gamma)}\)} For complex
systems in which individual system components represent the density of
some tangible quantity, it is relevant to consider the feasibility of
the system. Feasibilility assumes that the system is stable and values
of all components are positive at
equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Song2018}{21}}.
This is of particular interest for ecological communities because
population density cannot take negative values, meaning that ecological
systems need to be feasible for stability to be biologically
realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}}.
While my results are intended to be general to all complex systems, and
not restricted to species networks, I have also performed a feasibility
analysis on all matrices \(\mathbf{M}\) tested for stability, and
additionally for specific types of ecological
communities\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
(e.g., competitive, mutualist, predator-prey; see Supplementary
Information). Feasibility was unaffected by \(Var(\gamma)\) and instead
occurred with a fixed probability of \(1/2^{S}\), consistent with a
recent proof by Serván et
al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}} (see
Supplementary Information). Hence, for pure interacting species
networks, variation in component response rate (i.e., species generation
time) does not affect stability at biologically realistic species
densities.

@@

Here I have shown that the stability of large systems might often be
contigent upon variation in the response rates of their individual
components, meaning that factors such as rate of trait evolution (in
biological networks), transaction speed (in economic networks), or
communication speed (in social networks) need to be considered when
investigating the stability of complex systems. Variation in component
response rate becomes more likely to be critical for stability as system
size increases, and can ultimately increase the overall probability that
system stability is observed above that predicted by
May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}} classically
derived \(\sigma \sqrt{SC}\) criterion. The logic outlined here is
general, and potentially applies to any complex system in which
individual system components can vary in their reaction rates to system
perturbation.

@@

\[\mathbf{x^{*}} = -\left(\theta \mathbf{I} + (CS)^{-\delta}\mathbf{J} \right)^{-1}\mathbf{r}.\]

In the above, \(\mathbf{x^{*}}\) is the vector of species densities at
equilibrium. Feasibility is satisfied if all elements in
\(\mathbf{x^{*}}\) are positive. The matrix \(\mathbf{I}\) is the
identity matrix, and the value \(\theta\) is the strength of
intraspecific competition (diagonal elements). Diagonal values are set
to \(-1\), so \(\theta = -1\). The variable \(\delta\) is a
normalisation parameter that modulates the strength of interactions
(\(\sigma\)) for \(\mathbf{J}\). Implicitly, here \(\delta = 0\)
underlying strong interactions. Hence, \((CS)^{-\delta} = 1\), so in the
above, a diagonal matrix of -1s (\(\theta \mathbf{I}\)) is added to
\(\mathbf{J}\), which has a diagonal of all zeros and an off-diagonal
affecting species interactions (i.e., the expression \((CS)^{-\delta}\)
relates to May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}}
stability
criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}} by
\(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1\), and hence for my
purposes \((CS)^{-\delta} = 1\)). Given
\(\mathbf{A} = \theta\mathbf{I + J}\), the above criteria is therefore
reduced to the below (see
also\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}}),

\[\mathbf{x^{*} = -A^{-1}r}.\]

@@

\textbf{Figure 2: Distributions of eigenvalues before (a) and after (b)
introducing variation in component response rate
(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
system where \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\).
\textbf{a.} Eigenvalues plotted in the absence of \(Var(\gamma)\) where
\(E[\gamma] = 1\), versus \textbf{b.} eigenvalues plotted given
\(\gamma \sim \mathcal{U}(0, 2)\), which increases the variance of
interaction strengths (\(\sigma^{2}\)) but also creates a cluster of
eigenvalues toward the distribution's centre (-1, 0). Blue elipses in
both panels show the circle centred on the distribution in panel a.
Proportions of \(\Re(\lambda) < 0\) are 0.722 and 0.738 for a and b,
respectively.

>>>>

The stability of a complex system generally decreases with increasing
system size and interconnectivity, a counterintuitive result of
widespread importance across the physical, life, and social sciences.
Despite recent interest in the relationship between system properties
and stability, the effect of variation in the response rate of
individual system components remains unconsidered. Here I vary the
component response rates (\(\boldsymbol{\gamma}\)) of randomly generated
complex systems. I show that when component response rates vary, the
potential for system stability is markedly increased. Variation in
\(\boldsymbol{\gamma}\) becomes increasingly important for stability in
large but finite systems, where the probability of stability would
otherwise be negligible. At such extremes of simulated system size, the
largest stable complex systems would be unstable if not for
\(\boldsymbol{Var(\gamma)}\). My results therefore reveal a previously
unconsidered aspect of system stability that is likely to be pervasive
across all realistic complex systems.

@@

May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
stability criterion \(\sigma\sqrt{SC} < 1\) assumes that the expected
response rates (\(E[\gamma]\)) of individual components to perturbations
of the system are identical, but this is highly unlikely in any complex
system. In ecological communities, for example, the rate at which
population density changes following perturbation will depend on the
generation time of organisms, which might vary by orders of magnitude
among species. Species with short generation times will respond quickly
(high \(\gamma\)) to perturbations relative to species with long
generation times (low \(\gamma\)). Similarly, the speed at which
individual banks respond to perturbations in financial networks, or
individuals or institutions respond to perturbations in complex social
networks, is likely to vary. The effect of such variance on stability
has not been investigated in complex systems theory. Intuitively,
variation in \(\gamma\) might be expected to decrease system stability
by introducing a new source of variation into the system and thereby
increasing \(\sigma\). Here I show why, despite higher \(\sigma\),
realistic complex systems (such that \(S\) is high but finite) in which
\(\gamma\) varies are actually more likely to be stable.

@@

\textbf{Component response rates of random complex systems}. Rows in
\(\mathbf{M}\) define how a given component \(i\) is affected by other
components of the system, meaning that the rate of component response
time can be modelled by multiplying all row elements by a real scalar
value
\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
The distribution of \(\gamma\) over \(S\) components thereby models the
distribution of component response rates. An instructive example
compares one \(\mathbf{M}\) where \(\gamma_{i} = 1\) for all \(i\) in
\(S\) to the same \(\mathbf{M}\) when half of \(\gamma_{i} = 1.95\) and
half of \(\gamma_{i} = 0.05\). This models one system in which
\(\gamma\) is invariant and one in which \(\gamma\) varies, but systems
are otherwise identical (note that mean \(\gamma_{i} = 1\) in both
cases). I assume \(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\);
diagonal elements are set to \(-1\) and non-zero off-diagonal elements
are drawn randomly from \(\mathcal{N}(0, \sigma^{2})\). Rows are then
multiplied by \(\gamma_{i}\) to generate \(\mathbf{M}\). When
\(\gamma_{i} = 1\), eigenvalues of \(\mathbf{M}\) are distributed
uniformly within a circle centred at (\(-1, 0\)) with a radius of 1.265
(Fig. 1a). Hence, the real components of eigenvalues are highly unlikely
to all be negative when all \(\gamma_{i} = 1\). But when \(\gamma_{i}\)
values are separated into two groups, eigenvalues are no longer
uniformly distributed (Fig. 1b). Instead, two distinct clusters of
eigenvalues appear (red circles in Fig. 1b), one centred at
(\(-1.95, 0\)) and the other centred at (\(-0.05, 0\)). The former has a
large radius, but the real components have shifted to the left (in
comparison to when \(\gamma = 1\)), while the latter cluster has a
smaller radius, but real components have shifted to the right; all
\(\Re({\lambda}) < 0\). Overall, for 1 million randomly assembled
\(\mathbf{M}\), this division between slow and fast component response
rates results in more stable systems: 1 stable given \(\gamma = 1\)
versus 32 stable given \(\gamma = \{1.95, 0.5\}\).

Higher stability in systems with variation in \(\gamma\) can be observed
by sampling \(\gamma_{i}\) values from various distributions. I focus on
a uniform distribution where \(\gamma \sim \mathcal{U}(0, 2)\) (see
Supplementary Information for other distributions, which give similar
results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig. 1b), mean
\(\gamma = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
comparison of \(\mathbf{M}\) before and after the addition of variation
in component response rate. Figure 2 shows a comparison of eigenvalue
distributions given \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\). As
expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}}, when
\(\gamma = 1\), eigenvalues are distributed uniformly in a circle
centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 2.828.
Uniform variation in \(\gamma\) leads to a non-uniform distribution of
eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}--\protect\hyperlink{ref-Stone2017}{18}},
some of which are clustered locally near the centre of the distribution,
but others of which are spread outside the former radius of 2.828 (Fig
2b). The clustering and spreading of eigenvalues introduced by
\(Var(\gamma)\) can destabilise previously stable systems or stabilise
systems that are otherwise unstable. But where systems are otherwise too
complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
can often lead to stability above
May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
threshold \(\sigma\sqrt{SC} < 1\).

\textbf{Simulation of random \(\mathbf{M}\) across \(\mathbf{S}\)}. To
investigate the effect of \(Var(\gamma)\) on system stability, I
simulated random \(\mathbf{M}\) matrices at \(\sigma = 0.4\) and
\(C = 1\) across \(S = \{2, 3, ..., 49, 50\}\). One million
\(\mathbf{M}\) were simulated for each \(S\), and the stability of
\(\mathbf{M}\) was assessed given \(\gamma = 1\) versus
\(\gamma \sim \mathcal{U}(0, 2)\). For all \(S > 10\), I found that the
number of stable random systems was higher given \(Var(\gamma)\) than
when \(\gamma = 1\) (Fig. 3; see Supplementary Information for full
table of results), and that the difference between the probabilities of
observing a stable system increased with an increase in \(S\). In other
words, the potential for \(Var(\gamma)\) to drive stability increased
with system complexity and was most relevant for systems on the cusp of
being too complex to be realistically stable. For the highest values of
\(S\), nearly all systems that were stable given \(Var(\gamma)\) would
not have been stable given \(\gamma = 1\).

\textbf{Targeted manipulation of \(\mathbf{\gamma}\)}. To further
investigate the potential of \(Var(\gamma)\) to be stabilising, I used a
genetic algorithm. Genetic algorithms are heuristic tools that mimic
evolution by natural selection, and are useful when the space of
potential solutions (in this case, possible combinations of \(\gamma\)
values leading to stability in a large complex system) is too large to
search
exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}.
Generations of selection on \(\gamma\) value combinations to minimise
\(\max\left(\Re(\lambda)\right)\) demonstrated the potential for
\(Var(\gamma)\) to increase system stability. Across
\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
resulted in stable systems with probabilities that were up to four
orders of magnitude higher than when \(\gamma = 1\) (Fig. 4), meaning
that stability could often be achieved by manipulating \(S\) \(\gamma\)
values rather than \(S \times S\) \(\mathbf{M}\) elements (i.e., by
manipulating component response rates rather than interactions between
components).

\textbf{System feasibility given \(\mathbf{Var(\gamma)}\)} For complex
systems in which individual system components represent the density of
some tangible quantity, it is relevant to consider the feasibility of
the system. Feasibilility assumes that the system is stable and values
of all components are positive at
equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Song2018}{21}}.
This is of particular interest for ecological communities because
population density (\(N\)) cannot take negative values, meaning that
ecological systems need to be feasible for stability to be biologically
realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}}.
While my results are intended to be general to all complex systems, and
not restricted to species networks, I have also performed a feasibility
analysis on all matrices \(\mathbf{M}\) tested for stability, and
additionally for specific types of ecological
communities\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
(e.g., competitive, mutualist, predator-prey; see Supplementary
Information). I want to emphasise that \(\gamma\) is not interpreted as
population density in this analysis, but instead as a fundamental
property of species life history such as expected generation time.
Feasibility was unaffected by \(Var(\gamma)\) and instead occurred with
a fixed probability of \(1/2^{S}\), consistent with a recent proof by
Serván et al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}}
(see Supplementary Information). Hence, for pure interacting species
networks, variation in component response rate (i.e., species generation
time) does not affect stability at biologically realistic species
densities.

@@

Here I have shown that the stability of large systems might often be
contigent upon variation in the response rates of their individual
components, meaning that factors such as rate of trait evolution (in
biological networks), transaction speed (in economic networks), or
communication speed (in social networks) need to be considered when
investigating the stability of complex systems. Variation in component
response rate is more likely to be critical for stability in systems
that are especially large (but finite, see Gibbs et
al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17}}), and it can
ultimately increase the probability that system stability is observed
above that predicted by
May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}} classically
derived \(\sigma \sqrt{SC}\) criterion. The logic outlined here is
general, and potentially applies to any complex system in which
individual system components can vary in their reaction rates to system
perturbation.

@@

\[\mathbf{n^{*}} = -\left(\theta \mathbf{I} + (CS)^{-\delta}\mathbf{J} \right)^{-1}\mathbf{r}.\]

In the above, \(\mathbf{n^{*}}\) is the vector of species densities at
equilibrium. Feasibility is satisfied if all elements in
\(\mathbf{n^{*}}\) are positive. The matrix \(\mathbf{I}\) is the
identity matrix, and the value \(\theta\) is the strength of
intraspecific competition (diagonal elements). Diagonal values are set
to \(-1\), so \(\theta = -1\). The variable \(\delta\) is a
normalisation parameter that modulates the strength of interactions
(\(\sigma\)) for \(\mathbf{J}\). Implicitly, here \(\delta = 0\)
underlying strong interactions. Hence, \((CS)^{-\delta} = 1\), so in the
above, a diagonal matrix of -1s (\(\theta \mathbf{I}\)) is added to
\(\mathbf{J}\), which has a diagonal of all zeros and an off-diagonal
affecting species interactions (i.e., the expression \((CS)^{-\delta}\)
relates to May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}}
stability
criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}} by
\(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1\), and hence for my
purposes \((CS)^{-\delta} = 1\)). Given
\(\mathbf{A} = \theta\mathbf{I + J}\), the above criteria is therefore
reduced to the below (see
also\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}}),

\[\mathbf{n^{*} = -A^{-1}r}.\]

@@

\textbf{Figure 2: Distributions of eigenvalues before (a) and after (b)
introducing variation in component response rate
(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
system where \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\).
\textbf{a.} Eigenvalues plotted in the absence of \(Var(\gamma)\) where
\(E[\gamma] = 1\), versus \textbf{b.} eigenvalues plotted given
\(\gamma \sim \mathcal{U}(0, 2)\), which increases the variance of
interaction strengths (\(\sigma^{2}\)) but also creates a cluster of
eigenvalues toward the distribution's centre (-1, 0). Blue elipses in
both panels show the circle centred on the distribution in panel a.
Proportions of \(\Re(\lambda) < 0\) are 0.724 and 0.732 for a and b,
respectively.
